Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-04-30 doc/network.tex

\jd{Rewrite these equations a bit more consistently. What are we using for: index infection, current infection? $\tau$ is difference between them. Is it simpler just to use ' for dummies?}
\swp{I tried. $t$ is current time, $\tau$ is time difference. $s$ and $x$ are used for integrals.}
In a homogeneously mixing population, we can assume that the disease incidence $i(t)$ is the product of the current infectiousness of previously infected individuals and the proportion of the population susceptible, $S(t)$.
\begin{equation}
i(t) = S(t) \int K(s) i(t-s) ds = \RR \int g(s) i(t-s) ds,
\end{equation}
where $\RR = S(t) \RR_0$ is the effective reproductive number.
This model, also referred to as the renewal equation, can represent a wide range of epidemic models \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral, wallinga2007generation, roberts2007model}.
Over a period of time where the proportion of susceptible $S$ remains roughly constant, we would expect approximately exponential growth in incidence $i(t)$; assuming $i(t) = i(0) \exp(r t)$ yields the Euler-Lotka equation \citep{lotka1907relation}, which provides a direct link between speed and strength of an epidemic:
\begin{equation}
\frac{1}{\RR} = \int g(s) \exp(-r s) ds.
\end{equation}

@@

\jd{Can we just skip this \P, or put in an appendix or something?}
\swp{We decided to ask David for opinion}
\swp{I tried to rewrite the section to make things clearer; it seems to make this paragraph less boring and more relevant?}
On the other hand, the censored generation-interval distribution, $c_t(\tau)$, describes a distribution of \emph{all} infections that are $\tau$ time units apart from any cohort infection time $s$ before $t$.
Then, the censored generation-interval distribution is given by
\begin{equation}\label{eq:obsg}
c_t(\tau) = \frac{g(\tau) \int_0^t i(s-\tau) S(s) ds}{\int_0^t g(x) \int_0^t i(s-x) S(s) ds dx}
\end{equation}
Alternatively, the censored generation-interval distribution can be interpreted as a weighted distribution of the backward generation intervals:
\begin{equation}
c_t(\tau) \propto \int_0^t b_s(\tau) S(s) ds.
\end{equation}
Therefore, we expect the backward generation-interval distribution to be equivalent to the censored generation-interval distribution when $S \approx 1$, i.e., when the epidemic is growing exponentially;
however, this assumption quickly breaks down as the epidemic progresses (Figure~\ref{fig:censor}).

>>>>

\jd{Rewrite these equations a bit more consistently. What are we using for: index infection, current infection? $\tau$ is difference between them. Is it simpler just to use ' for dummies?}
\swp{I tried. $t$ is current time, $\tau$ is time difference. $s$ and $x$ are used for integrals.}
\jd{How about: $s$ as a dummy for $t$, $\sigma$ as a dummy for $\tau$?}

In a homogeneously mixing population, we can assume that the disease incidence $i(t)$ is the product of the current infectiousness of previously infected individuals and the proportion of the population susceptible, $S(t)$.
\begin{equation}
i(t) = S(t) \int K(s) i(t-s) ds = \RR \int g(s) i(t-s) ds,
\end{equation}
where $\RR = S(t) \RR_0$ is the effective reproductive number.
This model, also referred to as the renewal equation, can represent a wide range of epidemic models \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral, wallinga2007generation, roberts2007model}.
Over a period of time where the proportion of susceptible $S$ remains roughly constant, we would expect approximately exponential growth in incidence $i(t)$; assuming $i(t) = i(0) \exp(r t)$ yields the Euler-Lotka equation \citep{lotka1907relation}, which provides a direct link between speed and strength of an epidemic:
\begin{equation}
\frac{1}{\RR} = \int g(s) \exp(-r s) ds.
\end{equation}

@@

\jd{Can we just skip this \P, or put in an appendix or something?}
\swp{We decided to ask David for opinion}
\swp{I tried to rewrite the section to make things clearer; it seems to make this paragraph less boring and more relevant?}

On the other hand, the censored generation-interval distribution, $c_t(\tau)$, describes a distribution of \emph{all} infections that are $\tau$ time units apart from any cohort infection time $s$ before $t$.
Then, the censored generation-interval distribution is given by
\begin{equation}\label{eq:obsg}
c_t(\tau) = \frac{g(\tau) \int_0^t i(s-\tau) S(s) ds}{\int_0^t g(x) \int_0^t i(s-x) S(s) ds dx}
\end{equation}
Alternatively, the censored generation-interval distribution can be interpreted as a weighted distribution of the backward generation intervals:
\begin{equation}
c_t(\tau) \propto \int_0^t b_s(\tau) S(s) ds.
\end{equation}
Therefore, we expect the backward generation-interval distribution to be equivalent to the censored generation-interval distribution when $S \approx 1$, i.e., when the epidemic is growing exponentially;
however, this assumption quickly breaks down as the epidemic progresses (Figure~\ref{fig:censor}).
